Figures to insert in text

### 2018-04-15 Chapters/Chapter02.tex
The incredible animal behavior of Dr Linhares

@@

Psychological
miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue... how does this recall process work?

@@

Computational (fully experimental, with no voodoo-like mysterious properties)

@@

\section{The wasted effort of duplicated work}

As expected for such an important thery, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth. \citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}. Despite this, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it. Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate effort from researchers to build on top of previous work.

@@

    \item a 2013 Java implementation `Learning Intelligent Distribution Agent' (LIDA) developed by \citep{franklin_lida:_2014, snaider_integer_2013, snaider_modular_2014} Stan Franklin from the University of Memphis includes implementation. \footnote{http://ccrg.cs.memphis.edu/framework.html};

@@

Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas \citep{Anwar2003, snaider_integer_2013, snaider_modular_2014, franklin_lida:_2014}.  It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to Python\footnote{Python is sometimes called a `glue language'. That is, in my opinion, not the best metaphor.  A glue connects two things leaving an inflexible structure.  Python is perhaps best described as the interstate highway system of Programming; if something is out there, there is a way to reach it with Python.  In the comparison with Java, for instance, take the $popcnt(xor(b_i,b_j))$ operation, executed billions of times in SDM. How easy is it to program that for a particular GPU or FPGA with Java?}.  It is not available on an open-source repository.  It has a non-standard license, strange to the open-source community, \emph{the LIDA Framework Software NonExclusive, Non-Commercial Use License}.  Moreover, potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''... \emph{`before Memphis can accept it'} \footnote{What they are attempting to do with this bureaucracy remains unknown, the history of computing has not been kind to those who favored centralization \citep{ferguson_computer_2002}.}.

@@

We run the same simulation for a 256-bit SDM with 1,000,000 hard locations. The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}). Notice that the loss of 50\% of neurons barely affected the 256-bit SDM which remained functional even facing an enormous loss of 90\% of neurons.

@@

@@

% \subsection{Deep learning, multiple SDMs, and the animal behavior of Dr. Linhares}

% Deep neural networks, alphaGo, alphazero

\section{Illuminating, paving, clearing}

@@

% TODO

>>>>

% TODO GARY KLEIN !!!
GARY KLEIN!!!!

@@

% TODO Venn Diagram Psychology, Neuroscience, Cognitive modelling
% TODO Graph between these disciplines and how they might interact.

Psychological
orthogonality of concepts; life flashing before one's eyes; miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue... how does this recall process work?

@@

Cognitive modelling (fully experimental, with no voodoo-like mysterious properties)

@@

\section{The wasted effort of duplicated, AD-HOC, work}

As expected for such an important thery, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth. \citet{Linhares2011} has shown that SDM respects the limits of short-term memory \citep{Miller1955, Cowan2001}. Despite this, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it. Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate effort from researchers to build on top of previous work.

@@

    \item a 2013 Java implementation `Learning Intelligent Distribution Agent' (LIDA) developed by \citep{franklin_lida:_2014, snaider_integer_2013, snaider_modular_2014} Stan Franklin's group from the University of Memphis includes implementation. \footnote{http://ccrg.cs.memphis.edu/framework.html; see also http://ccrg.cs.memphis.edu/projects.html where they link to a github repository.};

@@

Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas \citep{Anwar2003, snaider_integer_2013, snaider_modular_2014, franklin_lida:_2014}.  It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to the combination Python and OpenCL proposed here\footnote{Python is sometimes called a `glue language'. That is, in my opinion, not the best metaphor.  A glue connects two things leaving an inflexible structure.  Python is perhaps best described as the interstate highway system of Programming; if something is out there, there is a way to reach it with Python.  In the comparison with Java, for instance, take the $popcnt(xor(b_i,b_j))$ operation, executed billions of times in SDM. How easy is it to program that for a particular GPU or FPGA with Java?}.  It has a non-standard license, strange to the open-source community, \emph{the LIDA Framework Software NonExclusive, Non-Commercial Use License}.  We have not found any parallelism in their code \citep{ccrg_ccrg_nodate}, which may make simulations slow or unfeasible. Moreover, potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''... \emph{`before Memphis can accept it'} \footnote{What they are attempting to do with this bureaucracy remains unclear, the history of computing has not been kind to those who favored centralization \citep{ferguson_computer_2002}.  We certainly refrain from contributing given the legal uncertainties of non-standard licenses and dubious processes --- even as we would like to link these libraries}.

@@

We run the same simulation for a 256-bit SDM with 1,000,000 hard locations. The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}). Notice that the loss of 50\% of neurons barely affected the 256-bit SDM... and it even remained functional when facing an enormous loss of 90\% of neurons!  What these results seem to imply is not only that the memory is extremely robust, but that a very small expected number of neurons activated in each access radius would preserve the most valueable information.  This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{Magic numbers}.  Is there a thing such as excessive robustness, and, if so, when does robustness become waste?\footnote{Consider, for instance, error-correcting codes in Information Theory or the analogous computer ECC memory in Electrical Engineering: tradeoffs between robustness and waste must be considered at each design decision; Why should it be different in the modeling of human memory?}

@@

% Post-defense TODO: Use the robustness of neurons as a guide.  A more quantitative study of this loss of neurons... between d=0 and d=critical_distance
% Post-defense TODO: magic numbers
% Post-defense TODO: mu - m sigma, in which the expected number of neurons is, say, 256 (instead of >1000).
% Post-defense TODO: study this.  The marginal increase of 10% of Hard locations increases by how many percentage points the critical distance? In fact, is there a way to study this algebrically?  That would be so cool!

\section{Flow versus flash:  blind driving; life flashing before one's eyes; and temperature}

NOT FOR THE THESIS; move this.  Lots of blah-blah here about psychologicallymodelling these things etc.

\begin{table}
\begin{tabular}{|c|c|}
\hline
English & My whole life flashed before my eyes\tabularnewline
\hline
\hline
Arabic & Wa mana shereet hayaty endance raayt al mawt beayny\tabularnewline
\hline
Dutch & Het leven aan je voorbij zien gaan\tabularnewline
\hline
French & J'ai vu toute ma vie défiler\tabularnewline
\hline
German & Mein ganzes Leben passierte Revue\tabularnewline
\hline
Italian & Mi é passata tutta la vita davanti agli occhi\tabularnewline
\hline
Latin Script & Zendegiam dar yek lahzeh az joloyeh Cheshmhayam gozasht\tabularnewline
\hline
Norwegian & Livet passerte (i revy) foran øynene mine\tabularnewline
\hline
Persian & \tabularnewline
\hline
Portuguese & A vida passou diante dos meus olhos\tabularnewline
\hline
Romanian & Mie mi-a trecut toata viata prin fata ochilor\tabularnewline
\hline
Russian & \tabularnewline
\hline
Spanish & Toda mi vida me pasó delante de los ojos\tabularnewline
\hline
Swedish & Livet passerade revy framför mina ögon\tabularnewline
\hline
\end{tabular}\caption{An expression that seems universal. Could it be that these are merely
conscious, psychological reports, of what one feels to have experienced?
Could this be explained by a theory of human memory?  It seems to be related to Hofstadter's temperature, here at truly high levels... Is it possible to simulate this?}
% TODO include persian and russian back, through inkscape pdf
% https://tex.stackexchange.com/a/370073/33832
\end{table}\label{table:lflashing}

% TODO look at previous paper on this and speculate it here in SDM, somewhere
% TODO include on book from the following:
%

\begin{quote}
.. body leaning in, until he just collapsed dead before completing his little circle in the grass. The NVA were right behind Sergeant Jefferson, coming through the hedgerow, and Power's life flashed before his eyes. He saw it all in a second, and he realized then and there that everything he had ever done, every decision he had ever made, it had all been to get him in this hole on this day. I'm dying' right here. That's it. It's over. Power shot an NVA who burst through the hedgerow with his ...
\end{quote}

Consider the above and that the expression on table \ref{table:lflashing} seem to be universal...

% https://books.google.com.br/books?id=OURuAAAAMAAJ&q=%22life+flashed+before+his+eyes%22+subject:%22History%22&dq=%22life+flashed+before+his+eyes%22+subject:%22History%22&hl=en&sa=X&ved=0ahUKEwjdr9aiu7faAhWHvJAKHdlGC7IQ6AEIXDAJ

% https://www.amazon.com/gp/offer-listing/044021310X/ref=olp_page_1?ie=UTF8&f_all=true
%

@@

% Post-Defense TODO:  SDM creating Voronoi Tesselations of concepts, with different sized areas for each.  This should solve i versus l.

\section{Deep learning, multiple SDMs, and the incredible animal behavior of Dr. Linhares}

In the name of all that is sacred, it is probably advisable to leave the following comments as originally made by my advisor, Dr. Linhares. To quote:

\begin{figure}[h!]
  \centering

  \subfloat[Bongard problem 71]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp071.png}}

  \subfloat[Bongard problem 72]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp072.png}}

  \subfloat[Bongard problem 73]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp073.png}}

  \caption{Bongard problems 71 - 73, courtesy of noted Bongardologist, Dr. Harry Foundalis.  What distinguishes the boxes on the left hand side to those on the right hand side?}
  \label{fig:bongard-problems-71}
\end{figure}

\begin{figure}[h!]
  \centering

  \subfloat[Bongard problem 74]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp074.png}}

  \subfloat[Bongard problem 75]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp075.png}}

  \subfloat[Bongard problem 76]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp076.png}}

  \label{fig:bongard-problems-74}
\end{figure}

\begin{figure}[h!]
  \centering

  \subfloat[Bongard problem 77]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp077.png}}

  \subfloat[Bongard problem 78]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp078.png}}

  \subfloat[Bongard problem 79]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp079.png}}

  \label{fig:bongard-problems-77}
\end{figure}

\begin{quote}

    Those deep neural networks... alphaGo... alphazero... very impressive, and to me very right and proper in a sense and also very wrong and misguided in another. I have not seriously studied the topic. Eric Nichols explained it to me in a napkin, and I watched some youtube videos and browsed the books. So what follows, like all great science, is completely based on gossip, hearsay, prejudice and anecdote.  Or at least that's how I believe all great \emph{hunches} over what is promising and what is a dead-end come from; that aesthetic sense that forms over time and seems to guide us towards what feels beautiful and correct and away from what seems foolish and wasteful...

    You know those animals that create those burrows underground, creatures like the prairie dogs?  There, underground, they build all sorts of passageways from here to where they want to go; and in the process they discover all that \emph{can't be passageway}.  What can't be passageway --- what can't be corridor --- will never be used as passageway, by definition.  And most of the world can't be passageway.  Think about the animal behavior: the beast won't give up! To me it's all very similar to what these systems seem to be doing... it seems they are molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...  You have these monstrous networks creating layers upon layers in which they differentiate and distribute the errors throughout all the nooks and cranies of, say, a domain like the game of Go.  Then, after some biblical number-crunching, it seems that there may be a very subtle way... however tortuous, however tight, however inprobable, however unconfortable.

    And these machines are exhibiting intelligence.  The victories in the game of Go seem to have shown that alphaGo did have vague, non-explicit, non-formal, but very strong intuitions about what the Go universe looks like.  So here is what I like and I find beautiful:  These gigantic clusters of inter-related micro impressions that the system has acquired and can relate to.  I imagine multiple SDMs exchanging information; gossiping about all the dead-ends out there and trying to find some tiny tortuous pathway, and it feels to me that that's a good and proper way to move forward in research.

    But it feels like a mistake to say that that is thinking.  It's not.  It is to me a great model of someone getting so attuned into an area that they are able to simply flow through tasks; in the positive psychology sense of the word flow.  It's very powerful and very intelligent:  but it is a form of intelligence without thinking.  Intelligence that can't make leaps from one arena to another; Intelligence that can't work on Bongard problems.  So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type... in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71, fig:bongard-problems-71, fig:bongard-problems-71}).

    Cognitive Science is, to me, the study of plausible interaction between system 1 and system 2.  We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa). We need both system 1 and system 2 to be plausible, in the psychological sense and in the neuroscientific sense.  AlphaGo, for instance, has this wonderful model of the game, reflecting many of the properties of system 1... but then falls back into silliness as soon as it turns to the other side of the problem:  it uses some montecarlo methods in very efficient fashion to go on to to the next probing of this wonderful system 1...

    Hofstadter has nailed the interactions and the content... analogies, fluid concepts, things seen through a different perspective, variations on a theme, etc.  Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.  Attaching something as beautiful as SDM to, say, some hyper-efficient alpha-beta pruning algorithm is nothing short of heresy against all the gods of science and must be punished with death without mercy. Seems like the most reasonable path going forward...

\end{quote}

\section{Not a real Conclusion}

@@

% TODO  Appendix?
